---
title: "class 08"
author: "Andy Kao"
date: "October 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Our first example with **kmeans()** function.

```{r}
#ex plot
plot(1:10,type="l")
```
back to kmeans
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
k <- kmeans(x, centers=2, nstart=20)
k
```
points in each cluster
```{r}
k$size
```
cluster assingment/membership
```{r}
k$cluster
table(k$cluster)
k$centers
```
plot
```{r}
plot(x,col=k$cluster)
points(k$centers, col="blue",pch=20,cex=3)
```

trying k n=3
```{r}
k3 <- kmeans(x, centers=3, nstart=20)
k3
k3$totss
k$totss
k3$tot.withinss
k$tot.withinss
```




hierarchical clustering
```{r}
# First we need to calculate point (dis)similarity
#   as the Euclidean distance between observations
dist_matrix <- dist(x) 
# The hclust() function returns a hierarchical 
#  clustering model
hc <- hclust(d = dist_matrix)
#hclust needs distance matrix imput
# the print method is not so useful here
hc  
#try plot
plot(hc)
abline(h=8,col="red")
cutree(hc, h=8)
cutree(hc,k=2)
```
try own section
```{r}
x <- rbind(
matrix(rnorm(100, mean=0, sd = 0.3),ncol=2),#c1
matrix(rnorm(100, mean = 1, sd=0.3),ncol=2),#c2
matrix(c(rnorm(50, mean = 1, sd = 0.3), #c3 
         rnorm(50, mean = 0, sd =0.3)),ncol=2))
colnames(x) <- c("x", "y")
plot(x)
col <- as.factor( rep(c("c1","c2","c3"), each=50) ) 
plot(x, col=col)
```
Q. Use the dist(), hclust(), plot() and cutree(
  functions to return 2 and 3 clusters 
 Q. How does this compare to your known 'col' groups?
```{r}
dist_matrix <- dist(x) 
hc <- hclust(d = dist_matrix)
plot(hc)
#return 2
hc2 <- cutree(hc, k=2)
hc2
#return 3
hc3 <- cutree(hc, k=3)
hc3

table(hc2)
table(hc3)
table(hc2,col)
table(hc3,col)
```
#pca
```{r}
 
## Initialize a blank 100 row by 10 column matrix 
## Initialize a blank 100 row by 10 column matrix 
mydata <- matrix(nrow=100, ncol=10) 
## Lets label the rows gene1, gene2 etc. to gene100 
rownames(mydata) <- paste("gene",  1:100, sep="") 
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5 
##   and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")  
colnames(mydata) <- c( paste("wt", 1:5, sep=""), paste("ko", 1:5, sep="") ) 
## Fill in some fake read counts 
for(i in 1:nrow(mydata)) { 
  wt.values <- rpois(5, lambda=sample
(x=10:1000, size=1)) 
  ko.values <- rpois(5, lambda=sample
(x=10:1000, size=1)) 
  mydata[i,] <- c(wt.values, ko.values) 
} 
head(mydata)
```
note: prcomp() funciton wants us to transpose our data so we flipped columns and rows
```{r}
head(t(mydata))
```

now we try pcrcomp()
```{r}
## lets do PCA 
pca <- prcomp(t(mydata), scale=TRUE) 
## See what is returned by the prcomp() function 
attributes(pca) 
```


```{r}
## lets do PCA 
pca <- prcomp(t(mydata), scale=TRUE) 
## A basic PC1 vs PC2 2-D plot 
plot(pca$x[,1], pca$x[,2]) 
## Precent variance is often more informative to look at 
pca.var <- pca$sdev^2 
pca.var.per <-round(pca.var/sum(pca.var)*100,1) 
pca.var.per
```



